---
title: "Week14_TAsession"
author: ""
date: "4/4/2020"
output: html_document
---


Packages needed.

```{r setup, include=FALSE}
library(tidyverse)
library(fpp2)

knitr::opts_chunk$set(echo = TRUE)
```

goog data set
* google stock data present in fpp2

```{r}
head(goog)
```
```{r}
# create training and validation of the Google stock data
goog.train <- window(goog, end = 900)
goog.test <- window(goog, start = 901)

```

Simple exponential smoothing
alpha: Value of smoothing parameter for the level
h: Number of periods for forecasting.
```{r}
ses.goog <- ses(goog.train, alpha = .2, h = 100)
autoplot(ses.goog)
```
Remove the trend

```{R}
goog.dif <- diff(goog.train) 
#goog.dif represents the change in stock price from the previous day
autoplot(goog.dif)
```
Differencing it effectively removes the rend from the data
```{R}
ses.goog.dif <- ses(goog.dif, alpha = .2, h = 100)
autoplot(ses.goog.dif)
```
```{r}
goog.dif.test <- diff(goog.test)
accuracy(ses.goog.dif, goog.dif.test)
```

Identifying optimal alpha parameter
```{R}
# identify optimal alpha parameter
alpha <- seq(.01, .99, by = .01)
RMSE <- NA
for(i in seq_along(alpha)) {
  fit <- ses(goog.dif, alpha = alpha[i], h = 100)
  RMSE[i] <- accuracy(fit, goog.dif.test)[2,2]
}

# convert to a data frame and idenitify min alpha value
alpha.fit <- data_frame(alpha, RMSE)
alpha.min <- filter(alpha.fit, RMSE == min(RMSE))

# plot RMSE vs. alpha
ggplot(alpha.fit, aes(alpha, RMSE)) +
  geom_line() +
  geom_point(data = alpha.min, aes(alpha, RMSE), size = 2, color = "blue")  
```
```{R}
# refit model with alpha = .05
ses.goog.opt <- ses(goog.dif, alpha = .01, h = 100)

# performance evaluations\
accuracy(ses.goog.opt, goog.dif.test)

# plotting results
p1 <- autoplot(ses.goog.opt) +
  theme(legend.position = "bottom")
p2 <- autoplot(goog.dif.test) +
  autolayer(ses.goog.opt, alpha = .5) +
  ggtitle("Predicted vs. actuals for the test data set")

gridExtra::grid.arrange(p1, p2, nrow = 1)
```
 *The predicted confidence intervals are narrower (left chart). 
 *The predicted versus actuals (right chart) you see that for most observations, our predicted confidence intervals did well.