

The dataset was collected from Airbnb with data on listings in the city of Asheville, NC. Here is the data provided for each listing:
•	room id: A unique number identifying an Airbnb listing.
•	host id: A unique number identifying an Airbnb host.
•	room type: One of 'Entire home/apt', 'Private room', or 'Shared room'
•	reviews: The number of reviews that a listing has received.
•	overall satisfaction: The average rating (out of five) that the listing has received from those visitors who left a review.
•	accommodates: The number of guests a listing can accommodate.
•	bedrooms: The number of bedrooms a listing offers.
•	price: The price (in USD) for a night stay. 

Loading the data
```{r}
# Load data and print head
airbnb_data <- read.csv("airbnb_data.csv",header = TRUE)
head(airbnb_data)
```

## Question 1: Fit a multiple linear regression model using price as the response variable and all others as predictor variables (Note: remove 'id' columns). Which variables are statistically significant in determining the price?


```{r}
# create model using 'lm' and print summary
model1 <- lm(price ~ room_type + reviews + overall_satisfaction + accommodates + bedrooms, data = airbnb_data)
summary(model1)
```
Statistically significant variables: overall_satisfaction, accommodates, bedrooms (can be concluded from p-values in the summary)

## Question 2: Interpret the coefficients for predictors: room type(Shared Room), bedrooms? 


Interpretations are as follows:
1) Room type(Shared Room): Holding all other variables constant, a listing for a shared room has an estimated price of 76.67 USD less than an entire home/apt. 
2) Bedrooms: Holding all other variables constant, the estimated price of a listing increases by 85.64 USD with an inremental bedroom in the property. 

## Question 3: Predict the price (nearest dollar) for a listing with the following factors: bedrooms = 1, accommodates = 2, reviews = 70, overall_satisfaction = 4, and room_type= 'Private room'.

```{r}
# prepare prediction data and use 'predict' to find the value
pred_data = data.frame(bedrooms = 1, accommodates = 2, reviews = 70, overall_satisfaction = 4, room_type = 'Private room')
predict(model1, pred_data)
```
The estimated price for such a listing is 66 dollars.

## Question 4: Identify outliers using Cook's distance approach. Remove points having Cook's distance > 1. Rerun the model after removal of these points and print summary.

```{r}
# Use cook's distance to identify outliers
cooks <-cooks.distance(model1)
which(cooks>1)
```
```{r}
# remove the outliers
airbnb_data_2 = airbnb_data[-c(94,95),]

# creating new model and print summary
model2 <- lm(price ~ room_type + reviews + overall_satisfaction + accommodates + bedrooms, data = airbnb_data_2)
summary(model2)
```
```{r}

```

